\documentclass{article}
\usepackage{graphicx} % For including diagrams

\title{MPI File Transfer Practical Work}
\author{Ha Tan Minh}
\date{}

\begin{document}

\maketitle

\section{Introduction}
The purpose of this practical is to implement a file transfer system using Message Passing Interface (MPI). The system involves multiple processes communicating and transferring files efficiently using MPI primitives.

\section{Protocol Design}
The file transfer protocol using MPI involves:
\begin{itemize}
    \item Dividing the file into chunks.
    \item Using one process (rank 0) as the master to distribute chunks to other processes.
    \item Collecting the chunks at the receiver side and reconstructing the file.
\end{itemize}


\section{Implementation}
The implementation involves an MPI-based Python script using the \texttt{mpi4py} library.

\subsection{MPI Code}
The MPI script handles both sending and receiving:
\begin{verbatim}
from mpi4py import MPI
import os

comm = MPI.COMM_WORLD
rank = comm.Get_rank()
size = comm.Get_size()

filename = "example.txt"
chunk_size = 1024  # Size of each chunk in bytes

if rank == 0:
    # Master process: read file and distribute chunks
    with open(filename, 'rb') as file:
        file_content = file.read()
    file_size = len(file_content)
    print(f"File size: {file_size} bytes")

    for i in range(1, size):
        start_index = (i - 1) * chunk_size
        end_index = min(start_index + chunk_size, file_size)
        comm.send(file_content[start_index:end_index], dest=i, tag=11)
    print("File chunks sent to workers.")

else:
    # Worker processes: receive chunks and save to file
    chunk = comm.recv(source=0, tag=11)
    with open(f"received_{rank}.txt", 'wb') as file:
        file.write(chunk)
    print(f"Process {rank} received and saved chunk.")

if rank == 0:
    # Master process: reassemble chunks at the receiver
    with open("received_file.txt", 'wb') as output_file:
        for i in range(1, size):
            with open(f"received_{i}.txt", 'rb') as input_file:
                output_file.write(input_file.read())
    print("File reassembled successfully.")
\end{verbatim}

\section{Results}
The MPI-based file transfer system was successfully tested. The following results were obtained:
\begin{itemize}
    \item File transferred: `example.txt`
    \item File size: 58 B
    \item Number of processes: 4
    \item Transfer time: ~1 second
\end{itemize}

\section{Roles}
Contributed to this project:
\begin{itemize}
    \item Ha Tan Minh: Developed the master-worker MPI code.
    \item Ha Tan Minh: Tested the file transfer with multiple processes.
    \item Ha Tan Minh: Prepared the report in LaTeX.
\end{itemize}

\end{document}
